\documentclass[11pt]{article}
\usepackage[utf8]{inputenc}
\usepackage[czech, english]{babel}
\usepackage[left=2cm,right=2cm,top=2cm,bottom=2cm]{geometry}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{amsmath}
\usepackage{tikz}
\usetikzlibrary{shapes.geometric}
\usepackage{hyperref}


% Itemize environment with small skips
\newenvironment{packeditemize}{
\begin{itemize}
  \setlength{\itemsep}{1pt}
  \setlength{\parskip}{0pt}
  \setlength{\parsep}{0pt}
}{\end{itemize}}


% Fancy footnote....
\usepackage{fancyhdr}
\pagestyle{fancy}
\usepackage{lastpage}
\rfoot{MATH 566 - 16, page \thepage/\pageref{LastPage}}
\cfoot{}
\rhead{}
\lhead{}
\renewcommand{\headrulewidth}{0pt}
\renewcommand{\footrulewidth}{0pt}


\newtheorem{theorem*}{Theorem} 

% Itemize environment with small skips
\newenvironment{packedenumerate}{
\begin{enumerate}
  \setlength{\itemsep}{1pt}
  \setlength{\parskip}{0pt}
  \setlength{\parsep}{0pt}
}{\end{enumerate}}


\newcounter{excounter}
\setcounter{excounter}{1}
\newcommand\question[2]{\vskip 1em  \noindent\textbf{\arabic{excounter}\addtocounter{excounter}{1}:} \emph{#1} \noindent#2}
\newcommand\solution[1]{\vskip 0.5em \noindent\textbf{Solution:} #1}
%\newcommand\like{\par \noindent\emph{(This question is: good - bad - ugly? Difficulty: 0-10:\hskip 3em )}} 

\newcommand\lecturer[1]{\textbf{#1}}
\newcommand\hideforstudent[1]{#1}


%\renewcommand\lecturer[1]{{\color{white} \textbf{#1} }}
%\renewcommand\hideforstudent[1]{{\color{white} #1 }}
%\renewcommand\solution[1]{{\color{white} \vskip 0.5em \noindent\textbf{Solution:} #1 }}



\setlength{\parindent}{0cm}
\setlength{\parskip}{0.1cm}

\begin{document}

Fall  2016, MATH-566

\centerline{{\Large \textbf{Network flows - Fast(er) Algorithm}}}

\textbf{Edmonds-Karp Algorithm}\\
\emph{Input:} Network $(G,u,s,t)$.\\
\emph{Output:} and $s$-$t$-flow $f$ of maximum value
\begin{enumerate}
\item $f(e) = 4$ for all $e \in E(G)$
\item while $f$-augmenting path exists:
\item \hskip 4em find shortest $f$-augmenting path $P$
\item \hskip 4em compute $\gamma := \min_{e \in E(P)} u_f(e)$
\item \hskip 4em augment $f$ along $P$ by $\gamma$ (as much as possible)
\end{enumerate}

Note that the shortest path can be implemented by \\
\textbf{BFS (Breath First Search) algorithm}:\\
\emph{Input:} Graph $G$, $s \in V(G)$.\\
\emph{Output:} spanning tree $T$ of shortest paths to $s$
\begin{enumerate}
\item $R=\{s\}$, $Q=(s)$, $T=(V,\emptyset)$.
\item while $Q$ is not empty:
\item \hskip 4em remove the first entry in $Q$, denote it by $u$.
\item \hskip 4em $\forall uv \in E(G)$, if $v \not\in R$ 
\item \hskip 8em             add $v$ at the end of $Q$;     add $v$ to $R$;  add $uv$ to $T$
\end{enumerate}

\question{}{
What is running time of $BFS$?
}
\solution{
$O(m)$. Every edge is touched at most twice.
}

\textbf{Lemma} 8.13
Let $f_1; f_2; \ldots$ be a sequence of flows such that $f_{i+1}$ results from $f_i$
by augmenting along $P_i$, where $P_i$ is a shortest $f_i$-augmenting path.
Then \\
(a) $|E(P_k)| \leq |E(P_{k+1})|$  for all $k$. \\
(b) $|E(P_k)|+2 \leq |E(P_l)|$ for all $k < l$ such that $P_k \cup P_l$ contains a pair of reverse
edges.
\question{}{
Prove (a). Consider edges $X$ of $P_k$ and $P_{k+1}$ (with multiplicity) together (and erase reverse edges).
Show that $|P_k|$ is at most half of the number of edges in $X$.
}
\solution{
Notice $X$ contains two edge disjoint paths since the outdegree of $s$ is 2, indegree of $t$ is 2 and all other vertices
are balanced. Notice that any path in $X$ was a candidate for $P_k$. Then 
\[
2|P_k| \leq |X| \leq |P_k| + |P_{k+1}|
\] 
}

\question{}{
Prove (b). Fix $k$ and consider the smallest $l > k$ such that $P_l$ uses a reverse edge of $P_k$. Use that there
was a reverse edge.
}
\solution{
Same as previous there was a reverse edge, so we can substract 2.
\[
2|P_k| \leq |X| \leq |P_k| + |P_{k+1}|-2
\] 
}

\question{}{
How many augmentations are needed in  Edmonds-Karp Algorithm?
What is the resulting running time?
}

\solution{
the length of the shortest path if at most $n$. In every augmenting path, at least one
edge is being saturated.
Every edge (or its reverse) is the saturated one in at most $\frac{n}{2}$ distances.
Together $\frac{mn}{2}$ iterations.

Every iteration takes one BFS, which takes $O(m)$. Hence the running time is $O(\frac{m^2n}{2})$.
}

\vskip 2em

\centerline{{\Large \textbf{Network flows as linear programs}}}

\question{}{
Formulate the maximum flow problem for network $(G,u,s,t)$ as a linear program $(P)$.
(Hint: Similar to shortest path.) Assume $G = (V,E)$.
}

\solution{
\[
(P) \begin{cases}
\text{maximize}  & \sum_{ut} f_{ut} - \sum_{tw} f_{tw}    \\
\text{subject to}  & \sum_{uv} f_{uv} -  \sum_{vw} f_{vw} = 0 \text{ for all } v \in V\setminus\{s,t\} \\
                           &  f_e \leq u(e) \text{ for all } e \in E\\
                           &  0 \leq f_e \text{ for all } e \in E
\end{cases}
\]
}

\question{}{
Write the dual $(D)$ to $(P)$. Use dual variables $y_v$, where $v\in V\setminus\{s,t\}$  for \quad $ \sum_{uv} f_{uv} -  \sum_{vw} f_{vw} = 0$, and
$z_e$ such that $e\in E$ for  $f_e\leq u(e)$.
}
\solution{
\[
(D) \begin{cases}
\text{minimize}  &  \sum_{e\in E} u(e)z_e \\
\text{subject to}
& -y_v+y_w+z_{vw}\geq 0 \text{ for all } vw\in E,\ v,w\in V\backslash\{s,t\}\\
& y_w+z_{sw}\geq 0 \text{ for all } sw\in E\\
& -y_v+z_{vs}\geq 0 \text{ for all } vs\in E\\
& -y_v+z_{vt}\geq 1\text{ for all }vt\in E\\
& y_w+z_{tw}\geq -1\text{ for all }tw\in E\\
& z_e\geq 0 \text{ for all } e\in E.
\end{cases}
\]
}
\question{}{
Add two artificial variables 
$y_s=0$ and $y_t=-1$. 
Then the constraints all unify to the form $-y_v+y_w+z_{vw}\geq 0$ for all $vw\in E$. Write the new program $(D')$.
}
\solution{
\[
(D') \begin{cases}
\text{minimize}  &  \sum_{e\in E} u(e)z_e \\
\text{subject to}
& -y_v+y_w+z_{vw}\geq 0 \text{ for all } vw\in E\\
& y_s = 0; y_t = -1\\
& z_e\geq 0 \text{ for all } e\in E.
\end{cases}
\]
Interpretation: every edge gives a bound how much of a decrease can occur.
Use the following figure to try to find a feasible solution (assign $z_e = 0$ and see why it is not a feasible 
solution.)
}

\question{}{
Recall that every $s$-$t$-flow can be decomposed into weighted $s$-$t$-paths.
Try to interpret $(D')$ using $s$-$t$ paths.
}
\solution{
\[
(D') \begin{cases}
\text{minimize}  &  \sum_{e\in E} u(e)z_e \\
\text{subject to}
& \sum_{e\in P}z_e \geq 1 \text{ for every } s\text{-}t-\text{path } P\\
& z_e\geq 0 \text{ for all } e\in E.
\end{cases}
\]
If $z_e$ is 0,1, it gives that every path must have some edge on it,
where $z_e = 1$ is an edge in a cut.
}

\end{document}






